\LARGE \bf Improving Automatic Whiteout++ through Feature Selection

### 2014-10-16 Paper/main.tex
Checkin items needed for submission

\section{Related Work}
The most relevant work done in this particular area was done with Automatic Whiteout++.  The main focus of this algorithm was to improve the original design to better detect and correct roll-on, roll-off, key repeat, and row substitution error.  Row subsitution errors account for $45\%$ of the total errors, so targeting and fixing these errors would produce a significant accurance boost\cite{clawson2008automatic}.  The algorithm utilizes decision trees in combination with Weka J48 to learn and classify keystrokes as errors or not.  Some of the common features used in this decision tree are keys pressed, timing information, a letter’s frequency in the English language, and adjacency to other keys.  There were also some newer features added to the algorithm, such as bi-letter/tri-letter frequencies, probability based on large plain-text corpus, and subsequent keypress evaluation.  This allowed for the additional ability to handle row substitution error, and to improve the accuracy of corrections on the original Automatic Whiteout.

@@

As Automatic Whiteout++, we used a set of experimental data that is an output of studies on two specific Dell and the Targus mini-QWERTY keyboards \cite{clawson2006mobile}. The study was carried out with 14 participants and 400 minutes of typing for each. The Twidor software was used to carry out the typing experiment. A total of 42,340 phrases were collected which accounted for 1,261,791 phrases. The data was further processed and each key press was characterised as correct or error. This set of data was used for training and testing.

@@

For each of these models, we report their performance on the training set averaged across ten folds (10-fold cross-validation) and then on the test set using the respective trained model in the following 6 tables. Each table has metrics for the four types of the off-by-one errors and for each type of error, we report the following metrics:
\begin{itemize}
\item TP: The number of actual errors that were correctly classified
\item FN: The number of actual errors which were incorrectly classified as some other type of error of nonerror
\item FP: The number of nonerrors that were classified as errors
\item TP + TN: Total number of actual errors
\item TP + FP: Total number of errors guessed
\item Precision: Percentage of total classification as errors that were actual errors
\item Recall: Percentage of total actual errors that were correctly classified
\end{itemize}

@@

From the perspective of cross validation metrics, we have a slight decrement (0.12\%) in precision and a slight enhancement (0.03\%) in recall for roll-on. For roll-off,  we have a slight decrement (0.21\%) in precision and a slight enhancement (0.62\%) in recall. For repeat, we perform almost the same with baseline in precision and recall. And for row-subsititute, we get a bit worse result (0.96\% less) in precision and (0.14\% less) recall.

@@

Compared against the test data metrics with only the division features, all the precisions suffer a hit and all the recalls except for repeats also perform worse. It's worthwile to note that repeats was the only error type which showed no performance gain with the division features as compared to the baseline.

Although the addition of the new features gave a better performance on the recall consistently, it negatively effected the precisions and thus calls for a trade-off.  Overall, the addition of the probability features improve performance over the baseline, but do not add any substantial improvement over the original division-based features and actually decrease some of the original improvements over the baseline.

>>>>

\section{Related Work}
The most relevant work done in this particular area was done with Automatic Whiteout++.  The main focus of this algorithm was to improve the original design to better detect and correct roll-on, roll-off, key repeat, and row substitution error.  Row substitution errors account for $45\%$ of the total errors, so targeting and fixing these errors would produce a significant accuracy boost\cite{clawson2008automatic}.  The algorithm utilizes decision trees in combination with Weka J48 to learn and classify keystrokes as errors or not.  Some of the common features used in this decision tree are keys pressed, timing information, a letter’s frequency in the English language, and adjacency to other keys.  There were also some newer features added to the algorithm, such as bi-letter/tri-letter frequencies, probability based on large plain-text corpus, and subsequent key press evaluation.  This allowed for the additional ability to handle row substitution error, and to improve the accuracy of corrections on the original Automatic Whiteout.

@@

As Automatic Whiteout++, we used a set of experimental data that is an output of studies on two specific Dell and the Targus mini-QWERTY keyboards \cite{clawson2006mobile}. The study was carried out with 14 participants and 400 minutes of typing for each. The Twidor software was used to carry out the typing experiment. A total of 42,340 phrases were collected which accounted for 1,261,791 phrases. The data was further processed and each key press was characterized as correct or error. This set of data was used for training and testing.

@@

For each of these models, we report their performance on the training set averaged across ten folds (10-fold cross-validation) and then on the test set using the respective trained model in the following 6 tables. Each table has metrics for the four types of the off-by-one errors and for each type of error, we report the following metrics:
\begin{itemize}
\item TP: The number of actual errors that were correctly classified
\item FN: The number of actual errors which were incorrectly classified as some other type of error of non-error
\item FP: The number of non-errors that were classified as errors
\item TP + TN: Total number of actual errors
\item TP + FP: Total number of errors guessed
\item Precision: Percentage of total classification as errors that were actual errors
\item Recall: Percentage of total actual errors that were correctly classified
\end{itemize}

@@

From the perspective of cross validation metrics, we have a slight decrease (0.12\%) in precision and a slight enhancement (0.03\%) in recall for roll-on. For roll-off,  we have a slight decrease (0.21\%) in precision and a slight enhancement (0.62\%) in recall. For repeat, we perform almost the same with baseline in precision and recall. And for row-substitute, we get a bit worse result (0.96\% less) in precision and (0.14\% less) recall.

@@

Comparing this test data against the test data metrics with only the division features, all the precisions suffer a hit and all the recalls except for repeats also perform worse. It's worthwhile to note that repeats was the only error type which showed no performance gain with the division features as compared to the baseline.

Although the addition of the new features gave a better performance on the recall consistently, it negatively effected the precision and thus calls for a trade-off.  Overall, the addition of the probability features improve performance over the baseline, but do not add any substantial improvement over the original division-based features and actually decrease some of the original improvements over the baseline.
